---
title: "RETRAUCI - Model Development and Validation"
output: html_document
date: "2026-07-27"
---

```{r}
library(dplyr)

archivo_cohorte <- "cohorte_retrascore.rds"
if (!file.exists(archivo_cohorte)) {
  stop(
    "cohorte_retrascore.rds was not found. ",
    "Run retrascore_validation_preprocessing.Rmd first."
  )
}
cohorte_retrascore <- readRDS(archivo_cohorte)
```

# 1. Analytical cohort from the paper and descriptive table

The 2015-2019 period is used: development in 2015-2017 and validation in
2018-2019. The eligible cohort requires a known in-hospital mortality outcome.
The analysis also requires a known 30-day outcome and complete data for all 20
candidate variables, without statistical imputation. The previous recoding of
empty MAIS values to zero is retained. The descriptive table and all models
use the same complete cases.

```{r}
candidatas_paper <- c(
  "edad_categoria", "genero_masculino", "precoag",
  "atencion_prehospitalaria_medica", "iot_prehospitalaria",
  "trauma_penetrante", "mecanismo_alto_riesgo", "pupilas",
  "gcs_menor_igual_8", "mais_cabeza", "mais_torax", "mais_abdomen",
  "mais_ext_superior", "mais_ext_inferior", "mais_externo",
  "fallo_hemodinamico", "fallo_respiratorio", "coagulopatia_24h",
  "ventilacion_mecanica", "hemorragia_masiva"
)
variables_regresion <- c("mortalidad_30d", "año_trauma", candidatas_paper)
cohorte_paper <- cohorte_retrascore %>%
  filter(año_trauma >= 2015, año_trauma <= 2019)
casos_completos_paper <- complete.cases(cohorte_paper[variables_regresion])
cohorte_analitica_paper <- cohorte_paper[casos_completos_paper, ]

flujo_cohorte_paper <- do.call(rbind, lapply(
  list(Derivacion = 2015:2017, Validacion = 2018:2019),
  function(anios) {
    d <- cohorte_paper[cohorte_paper$año_trauma %in% anios, ]
    completo <- complete.cases(d[variables_regresion])
    data.frame(
      periodo = paste(range(anios), collapse = "-"),
      elegibles = nrow(d),
      outcome_ausente = sum(is.na(d$mortalidad_30d)),
      excluidos_por_cualquier_ausente = sum(!completo),
      casos_completos = sum(completo)
    )
  }
))
rownames(flujo_cohorte_paper) <- NULL
flujo_cohorte_paper

datos_tabla_1 <- cohorte_analitica_paper

tabla_1_derivacion <- datos_tabla_1 %>%
  filter(año_trauma >= 2015, año_trauma <= 2017)

tabla_1_validacion <- datos_tabla_1 %>%
  filter(año_trauma >= 2018, año_trauma <= 2019)

grupos_tabla_1 <- list(
  All = datos_tabla_1,
  Derivation = tabla_1_derivacion,
  Validation = tabla_1_validacion
)

formatear_mediana_ric <- function(x) {
  x <- x[!is.na(x)]
  if (length(x) == 0) return(NA_character_)

  cuantiles <- quantile(x, c(0.25, 0.50, 0.75), names = FALSE)
  sprintf(
    "%g (%g-%g)",
    cuantiles[2], cuantiles[1], cuantiles[3]
  )
}

formatear_n_porcentaje <- function(x, valor = 1) {
  denominador <- sum(!is.na(x))
  if (denominador == 0) return(NA_character_)

  numerador <- sum(x == valor, na.rm = TRUE)
  sprintf("%.1f", 100 * numerador / denominador)
}

fila_continua_tabla_1 <- function(variable, etiqueta) {
  valores <- vapply(
    grupos_tabla_1,
    function(datos) formatear_mediana_ric(datos[[variable]]),
    character(1)
  )
  c(Variable = etiqueta, valores)
}

fila_binaria_tabla_1 <- function(variable, etiqueta, valor = 1) {
  valores <- vapply(
    grupos_tabla_1,
    function(datos) formatear_n_porcentaje(datos[[variable]], valor),
    character(1)
  )
  c(Variable = etiqueta, valores)
}

fila_encabezado_tabla_1 <- function(etiqueta) {
  c(Variable = etiqueta, All = "", Derivation = "", Validation = "")
}

filas_tabla_1 <- rbind(
  fila_continua_tabla_1("edad", "Age (years)"),
  fila_encabezado_tabla_1("Age (groups)"),
  fila_binaria_tabla_1("edad_categoria", "  < 50", "<50"),
  fila_binaria_tabla_1("edad_categoria", "  50-65", "50-65"),
  fila_binaria_tabla_1("edad_categoria", "  66-75", "66-75"),
  fila_binaria_tabla_1("edad_categoria", "  > 75", ">75"),
  fila_binaria_tabla_1("genero_masculino", "Sex (% male)"),
  fila_binaria_tabla_1("precoag", "PRECOAG"),
  fila_binaria_tabla_1(
    "atencion_prehospitalaria_medica",
    "Pre-hospital attention"
  ),
  fila_binaria_tabla_1(
    "iot_prehospitalaria",
    "Pre-hospital intubation"
  ),
  fila_encabezado_tabla_1("Type of injury"),
  fila_binaria_tabla_1(
    "tipo_lesion", "  Road Traffic accident", "Road Traffic accident"
  ),
  fila_binaria_tabla_1("tipo_lesion", "  Fall", "Fall"),
  fila_binaria_tabla_1(
    "tipo_lesion", "  Occupational accident", "Occupational accident"
  ),
  fila_binaria_tabla_1(
    "tipo_lesion", "  Sport related", "Sport related"
  ),
  fila_binaria_tabla_1("tipo_lesion", "  Assault", "Assault"),
  fila_binaria_tabla_1(
    "tipo_lesion", "  Self-injury", "Self-injury"
  ),
  fila_binaria_tabla_1("tipo_lesion", "  Unknown", "Unknown"),
  fila_binaria_tabla_1("tipo_lesion", "  Other", "Other"),
  fila_binaria_tabla_1(
    "mecanismo_alto_riesgo",
    "High-risk mechanism"
  ),
  fila_binaria_tabla_1("trauma_penetrante", "Mechanism penetrant"),
  fila_encabezado_tabla_1("PUPILS"),
  fila_binaria_tabla_1("pupilas", "  Normal", "normal"),
  fila_binaria_tabla_1(
    "pupilas", "  Unilateral mydriasis", "midriasis unilateral"
  ),
  fila_binaria_tabla_1(
    "pupilas", "  Bilateral mydriasis", "midriasis bilateral"
  ),
  fila_continua_tabla_1("gcs_total", "GCS"),
  fila_binaria_tabla_1("gcs_menor_igual_8", "GCS <= 8 (%)"),
  fila_binaria_tabla_1("mais_cabeza", "MAIS-Head"),
  fila_binaria_tabla_1("mais_torax", "MAIS-Thorax"),
  fila_binaria_tabla_1("mais_abdomen", "MAIS-Abdomen"),
  fila_binaria_tabla_1("mais_ext_superior", "MAIS-Ext Upper"),
  fila_binaria_tabla_1("mais_ext_inferior", "MAIS-Ext Lower"),
  fila_binaria_tabla_1("mais_externo", "MAIS-External"),
  fila_binaria_tabla_1(
    "fallo_hemodinamico", "Haemodynamic failure"
  ),
  fila_binaria_tabla_1("fallo_respiratorio", "Respiratory failure"),
  fila_binaria_tabla_1("coagulopatia_24h", "Coagulopathy"),
  fila_binaria_tabla_1(
    "ventilacion_mecanica", "Mechanical ventilation"
  ),
  fila_binaria_tabla_1("hemorragia_masiva", "Massive haemorrhage"),
  fila_continua_tabla_1("dias_estancia_uci", "ICU LOS (days)"),
  fila_binaria_tabla_1(
    "mortalidad_intrahospitalaria", "In-hospital mortality"
  ),
  fila_binaria_tabla_1("mortalidad_30d", "30-day mortality")
)

tabla_caracteristicas_cohortes <- as.data.frame(
  filas_tabla_1,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

names(tabla_caracteristicas_cohortes) <- c(
  "Variable",
  paste0("All\nN = ", nrow(datos_tabla_1)),
  paste0("Derivation\nN = ", nrow(tabla_1_derivacion)),
  paste0("Validation\nN = ", nrow(tabla_1_validacion))
)

tabla_caracteristicas_cohortes
```

# 2. LASSO selection in the development sample

The three requested steps are followed: temporal cohort definition, LASSO over
20 candidate variables, and unpenalized logistic regression after selection.
Source: [RETRASCORE paper](https://link.springer.com/article/10.1186/s13054-021-03845-6).

Reproducibility decisions made here: 10 outcome-stratified folds, seed 20260903,
predictor standardization, and lambda.1se selection based on binomial deviance.
lambda.min is also reported, without choosing between the two based on the
2018-2019 validation sample. These settings are not specified in the main text
of the paper. If LASSO retains any age or pupil category, the hierarchy
principle is applied and the complete categorical variable is retained in the
subsequent regression. Therefore, a zero coefficient for a particular level
does not mean that the level is removed from the final model when another level
of the same variable has been selected. Reference categories are shown
explicitly in the tables even though they do not have their own coefficient.
Selection of the 13 published variables is not forced.

The current coding is retained: MAIS from mais1-mais6, empty MAIS values coded
as zero, and other/unknown grouped as high risk at the user's request. This last
grouping is not the literal published definition. Deaths after 30 days are
treated as 30-day survivors. Transfers with an unknown hospital outcome remain
excluded. The 100 random splits, 1,000 bootstrap samples, and shrinkage
adjustment used in the paper have not been added at this stage; the following
coefficients are from the post-LASSO regression without that adjustment.

```{r}
if (!requireNamespace("glmnet", quietly = TRUE)) {
  stop("Install glmnet in RStudio before running this block.")
}
datos_regresion <- cohorte_analitica_paper
datos_derivacion <- datos_regresion %>%
  filter(año_trauma >= 2015, año_trauma <= 2017)
datos_validacion <- datos_regresion %>%
  filter(año_trauma >= 2018, año_trauma <= 2019)
stopifnot(nrow(datos_derivacion) > 0, nrow(datos_validacion) > 0)
stopifnot(length(unique(datos_derivacion$mortalidad_30d)) == 2,
          length(unique(datos_validacion$mortalidad_30d)) == 2)

formula_candidatas <- reformulate(candidatas_paper)
matriz_diseno <- model.matrix(formula_candidatas, data = datos_derivacion)
indices_terminos <- attr(matriz_diseno, "assign")[-1]
terminos_candidatos <- attr(terms(formula_candidatas), "term.labels")
x_lasso <- matriz_diseno[, -1, drop = FALSE]
y_lasso <- datos_derivacion$mortalidad_30d

set.seed(20260903)
foldid <- integer(length(y_lasso))
for (clase in c(0, 1)) {
  indices <- which(y_lasso == clase)
  stopifnot(length(indices) >= 10)
  foldid[indices] <- sample(rep(1:10, length.out = length(indices)))
}
cv_lasso <- glmnet::cv.glmnet(
  x = x_lasso, y = y_lasso, family = "binomial", alpha = 1,
  foldid = foldid, type.measure = "deviance", standardize = TRUE
)
criterio_lambda <- "lambda.1se"
beta_lasso <- as.matrix(coef(cv_lasso, s = criterio_lambda))[-1, 1]
beta_lasso_min <- as.matrix(coef(cv_lasso, s = "lambda.min"))[-1, 1]
tabla_seleccion_lasso <- data.frame(
  variable = terminos_candidatos[indices_terminos],
  termino = colnames(x_lasso),
  beta_lasso = as.numeric(beta_lasso),
  coeficiente_no_cero_lasso = as.numeric(beta_lasso) != 0,
  beta_lambda_min = as.numeric(beta_lasso_min),
  es_referencia = FALSE,
  orden = seq_along(beta_lasso) * 2,
  row.names = NULL
)
predictores_regresion <- unique(
  tabla_seleccion_lasso$variable[
    tabla_seleccion_lasso$coeficiente_no_cero_lasso
  ]
)

# The final decision is made at the variable level, not by isolated category.
tabla_seleccion_lasso$variable_seleccionada <-
  tabla_seleccion_lasso$variable %in% predictores_regresion
tabla_seleccion_lasso$incluido_modelo_final <-
  tabla_seleccion_lasso$variable_seleccionada

# model.matrix omits reference categories. They are added only to document
# that they belong to the factor; beta_lasso does not apply to these rows.
variables_con_referencia <- c("edad_categoria", "pupilas")
terminos_referencia <- c("edad_categoria<50", "pupilasnormal")
orden_referencia <- vapply(
  variables_con_referencia,
  function(v) min(tabla_seleccion_lasso$orden[
    tabla_seleccion_lasso$variable == v
  ]) - 1,
  numeric(1)
)
filas_referencia_lasso <- data.frame(
  variable = variables_con_referencia,
  termino = terminos_referencia,
  beta_lasso = NA_real_,
  coeficiente_no_cero_lasso = NA,
  beta_lambda_min = NA_real_,
  es_referencia = TRUE,
  orden = orden_referencia,
  variable_seleccionada = variables_con_referencia %in%
    predictores_regresion,
  incluido_modelo_final = variables_con_referencia %in%
    predictores_regresion,
  row.names = NULL
)
tabla_seleccion_lasso <- rbind(
  tabla_seleccion_lasso,
  filas_referencia_lasso
)
tabla_seleccion_lasso <- tabla_seleccion_lasso[
  order(tabla_seleccion_lasso$orden),
  setdiff(names(tabla_seleccion_lasso), "orden")
]
rownames(tabla_seleccion_lasso) <- NULL
resumen_lasso <- data.frame(
  n_derivacion = nrow(datos_derivacion),
  candidatas = length(candidatas_paper),
  columnas_diseno = ncol(x_lasso),
  lambda_min = cv_lasso$lambda.min,
  lambda_1se = cv_lasso$lambda.1se,
  variables_seleccionadas = length(predictores_regresion)
)
resumen_lasso
tabla_seleccion_lasso
```

# 3. Logistic regression after LASSO

The coefficients of the selected variables are re-estimated exclusively in the
development sample. The same patients used for LASSO are retained even when a
candidate variable has been discarded. The validation sample is not involved
in selection or model fitting.

```{r}
formula_regresion <- if (length(predictores_regresion)) {
  reformulate(predictores_regresion, response = "mortalidad_30d")
} else {
  mortalidad_30d ~ 1
}

formula_regresion <- mortalidad_30d ~
  edad_categoria +
  precoag +
  iot_prehospitalaria +
  mecanismo_alto_riesgo +
  pupilas +
  gcs_menor_igual_8 +
  mais_cabeza +
  mais_torax +
  fallo_hemodinamico +
  fallo_respiratorio +
  coagulopatia_24h +
  ventilacion_mecanica +
  hemorragia_masiva

# Keep subsequent tables synchronized with any manual change made to
# formula_regresion.
predictores_modelo_final <- attr(
  terms(formula_regresion),
  "term.labels"
)

modelo_regresion_logistica <- glm(
  formula_regresion,
  data = datos_derivacion,
  family = binomial(link = "logit")
)
stopifnot(modelo_regresion_logistica$converged)
if (anyNA(coef(modelo_regresion_logistica))) {
  stop("Non-estimable coefficients: check collinearity before validation.")
}

# Model evaluation in development and temporal validation -----------------

calcular_auc <- function(outcome, probabilidad) {
  casos <- outcome == 1
  controles <- outcome == 0
  n_casos <- sum(casos)
  n_controles <- sum(controles)

  if (n_casos == 0 || n_controles == 0) return(NA_real_)

  rangos <- rank(probabilidad, ties.method = "average")
  (sum(rangos[casos]) - n_casos * (n_casos + 1) / 2) /
    (n_casos * n_controles)
}

evaluar_modelo <- function(modelo, datos, muestra) {
  probabilidad <- predict(modelo, newdata = datos, type = "response")
  predictor_lineal <- predict(modelo, newdata = datos, type = "link")

  modelo_intercepto <- glm(
    mortalidad_30d ~ 1,
    offset = predictor_lineal,
    data = datos,
    family = binomial(link = "logit")
  )
  modelo_pendiente <- glm(
    mortalidad_30d ~ predictor_lineal,
    data = datos,
    family = binomial(link = "logit")
  )

  data.frame(
    muestra = muestra,
    n = nrow(datos),
    muertes_30d = sum(datos$mortalidad_30d),
    auc = calcular_auc(datos$mortalidad_30d, probabilidad),
    brier = mean((datos$mortalidad_30d - probabilidad)^2),
    intercepto_calibracion = coef(modelo_intercepto)[1],
    pendiente_calibracion = coef(modelo_pendiente)[2],
    row.names = NULL
  )
}

tabla_rendimiento_modelo <- rbind(
  evaluar_modelo(
    modelo_regresion_logistica,
    datos_derivacion,
    "Derivación 2015-2017"
  ),
  evaluar_modelo(
    modelo_regresion_logistica,
    datos_validacion,
    "Validación 2018-2019"
  )
)

tabla_rendimiento_modelo[, c(
  "auc", "brier", "intercepto_calibracion", "pendiente_calibracion"
)] <- round(
  tabla_rendimiento_modelo[, c(
    "auc", "brier", "intercepto_calibracion", "pendiente_calibracion"
  )],
  3
)

coeficientes_regresion <- summary(
  modelo_regresion_logistica
)$coefficients

# Dynamic table: only terms actually included in the model.
# Conventional Wald CIs; they do not include LASSO selection uncertainty.
tabla_regresion_logistica <- data.frame(
  Variable = rownames(coeficientes_regresion),
  beta = coeficientes_regresion[, "Estimate"],
  Puntos = ifelse(
    rownames(coeficientes_regresion) == "(Intercept)",
    NA_integer_,
    as.integer(round(2 * coeficientes_regresion[, "Estimate"]))
  ),
  OR = exp(coeficientes_regresion[, "Estimate"]),
  IC95_inferior = exp(coeficientes_regresion[, "Estimate"] -
    1.96 * coeficientes_regresion[, "Std. Error"]),
  IC95_superior = exp(coeficientes_regresion[, "Estimate"] +
    1.96 * coeficientes_regresion[, "Std. Error"]),
  orden = seq_len(nrow(coeficientes_regresion)) * 2,
  row.names = NULL
)

# Reference categories belong to the model, but R represents them with
# beta = 0, OR = 1, and 0 points, without estimating a separate CI.
referencias_modelo <- data.frame(
  variable_padre = c("edad_categoria", "pupilas"),
  Variable = c("edad_categoria<50", "pupilasnormal"),
  stringsAsFactors = FALSE
)
referencias_modelo <- referencias_modelo[
  referencias_modelo$variable_padre %in% predictores_modelo_final,
  , drop = FALSE
]
filas_referencia_regresion <- do.call(rbind, lapply(
  seq_len(nrow(referencias_modelo)),
  function(i) {
    variable_padre <- referencias_modelo$variable_padre[i]
    posiciones <- grep(
      paste0("^", variable_padre),
      tabla_regresion_logistica$Variable
    )
    data.frame(
      Variable = referencias_modelo$Variable[i],
      beta = 0,
      Puntos = 0L,
      OR = 1,
      IC95_inferior = NA_real_,
      IC95_superior = NA_real_,
      orden = min(tabla_regresion_logistica$orden[posiciones]) - 1,
      row.names = NULL
    )
  }
))
if (nrow(referencias_modelo) > 0) {
  tabla_regresion_logistica <- rbind(
    tabla_regresion_logistica,
    filas_referencia_regresion
  )
}
tabla_regresion_logistica <- tabla_regresion_logistica[
  order(tabla_regresion_logistica$orden),
  setdiff(names(tabla_regresion_logistica), "orden")
]
rownames(tabla_regresion_logistica) <- NULL

resumen_muestras_regresion <- data.frame(
  muestra = c("Derivación 2015-2017", "Validación 2018-2019"),
  n = c(nrow(datos_derivacion), nrow(datos_validacion)),
  muertes_30d = c(
    sum(datos_derivacion$mortalidad_30d),
    sum(datos_validacion$mortalidad_30d)
  ),
  mortalidad_pct = round(
    100 * c(
      mean(datos_derivacion$mortalidad_30d),
      mean(datos_validacion$mortalidad_30d)
    ),
    2
  )
)

resumen_muestras_regresion
tabla_regresion_logistica
tabla_rendimiento_modelo
```

# 4. Logistic regression using the variables selected by LASSO

This additional model is fitted only with the variables selected by LASSO at
`lambda.1se`. Unlike the regression in Section 3, the predictor formula is not
manually specified. It is built directly from `predictores_regresion`. The
model is fitted in the derivation sample and evaluated without refitting in the
temporal validation sample.

```{r}
# Build the regression formula directly from the LASSO-selected variables.
formula_regresion_seleccion_lasso <- if (length(predictores_regresion) > 0) {
  reformulate(predictores_regresion, response = "mortalidad_30d")
} else {
  mortalidad_30d ~ 1
}

variables_modelo_seleccion_lasso <- data.frame(
  variable = predictores_regresion,
  row.names = NULL
)

modelo_regresion_seleccion_lasso <- glm(
  formula_regresion_seleccion_lasso,
  data = datos_derivacion,
  family = binomial(link = "logit")
)

stopifnot(modelo_regresion_seleccion_lasso$converged)
stopifnot(nobs(modelo_regresion_seleccion_lasso) == nrow(datos_derivacion))
if (anyNA(coef(modelo_regresion_seleccion_lasso))) {
  stop(
    "Non-estimable coefficients in the LASSO-selected regression: ",
    "check collinearity before validation."
  )
}

# Estimate performance in derivation and temporal validation without refitting.
tabla_rendimiento_seleccion_lasso <- rbind(
  evaluar_modelo(
    modelo_regresion_seleccion_lasso,
    datos_derivacion,
    "Derivación 2015-2017"
  ),
  evaluar_modelo(
    modelo_regresion_seleccion_lasso,
    datos_validacion,
    "Validación 2018-2019"
  )
)

columnas_redondeo_lasso <- c(
  "auc", "brier", "intercepto_calibracion", "pendiente_calibracion"
)
tabla_rendimiento_seleccion_lasso[, columnas_redondeo_lasso] <- round(
  tabla_rendimiento_seleccion_lasso[, columnas_redondeo_lasso],
  3
)

# Create the coefficient table for the new regression.
coeficientes_seleccion_lasso <- summary(
  modelo_regresion_seleccion_lasso
)$coefficients

tabla_regresion_seleccion_lasso <- data.frame(
  Variable = rownames(coeficientes_seleccion_lasso),
  beta = coeficientes_seleccion_lasso[, "Estimate"],
  Puntos = ifelse(
    rownames(coeficientes_seleccion_lasso) == "(Intercept)",
    NA_integer_,
    as.integer(round(2 * coeficientes_seleccion_lasso[, "Estimate"]))
  ),
  OR = exp(coeficientes_seleccion_lasso[, "Estimate"]),
  IC95_inferior = exp(
    coeficientes_seleccion_lasso[, "Estimate"] -
      1.96 * coeficientes_seleccion_lasso[, "Std. Error"]
  ),
  IC95_superior = exp(
    coeficientes_seleccion_lasso[, "Estimate"] +
      1.96 * coeficientes_seleccion_lasso[, "Std. Error"]
  ),
  orden = seq_len(nrow(coeficientes_seleccion_lasso)) * 2,
  row.names = NULL
)

# Add the reference categories for selected categorical variables.
referencias_seleccion_lasso <- data.frame(
  variable_padre = c("edad_categoria", "pupilas"),
  Variable = c("edad_categoria<50", "pupilasnormal"),
  stringsAsFactors = FALSE
)
referencias_seleccion_lasso <- referencias_seleccion_lasso[
  referencias_seleccion_lasso$variable_padre %in% predictores_regresion,
  , drop = FALSE
]

if (nrow(referencias_seleccion_lasso) > 0) {
  filas_referencia_seleccion_lasso <- do.call(rbind, lapply(
    seq_len(nrow(referencias_seleccion_lasso)),
    function(i) {
      variable_padre <- referencias_seleccion_lasso$variable_padre[i]
      posiciones <- grep(
        paste0("^", variable_padre),
        tabla_regresion_seleccion_lasso$Variable
      )
      data.frame(
        Variable = referencias_seleccion_lasso$Variable[i],
        beta = 0,
        Puntos = 0L,
        OR = 1,
        IC95_inferior = NA_real_,
        IC95_superior = NA_real_,
        orden = min(tabla_regresion_seleccion_lasso$orden[posiciones]) - 1,
        row.names = NULL
      )
    }
  ))

  tabla_regresion_seleccion_lasso <- rbind(
    tabla_regresion_seleccion_lasso,
    filas_referencia_seleccion_lasso
  )
}

tabla_regresion_seleccion_lasso <- tabla_regresion_seleccion_lasso[
  order(tabla_regresion_seleccion_lasso$orden),
  setdiff(names(tabla_regresion_seleccion_lasso), "orden")
]
rownames(tabla_regresion_seleccion_lasso) <- NULL

formula_regresion_seleccion_lasso
variables_modelo_seleccion_lasso
tabla_regresion_seleccion_lasso
tabla_rendimiento_seleccion_lasso
```
