---
title: "RETRAUCI - Preprocessing"
output: html_document
date: "2026-07-27"
---

```{r}
library(dplyr)
library(lubridate)
```

```{r}
df <- read.csv2(
  "C:/Users/user/OneDrive - udl.cat/Desktop/RETRASCORE_VALIDATION/retrauci.csv",
  stringsAsFactors = TRUE
)

#summary(df)
```

```{r}
# ============================================================
# TABLE OF EMPTY CELLS, LITERAL "NAS" VALUES, AND VALID VALUES
# FOR ALL VARIABLES
# ============================================================

resumen_ausentes <- do.call(
  rbind,
  lapply(names(df), function(variable) {

    # Temporarily convert to text without modifying the original data frame
    x <- trimws(as.character(df[[variable]]))

    n_total <- length(x)

    # Empty Excel cells are usually imported as NA.
    # This also includes empty strings or values containing only spaces.
    es_vacio <- is.na(x) | (!is.na(x) & x == "")

    # Literal "NAS" text, case-insensitive
    es_nas <- !is.na(x) & toupper(x) == "NAS"

    # Values that are neither empty nor "NAS"
    es_valido <- !es_vacio & !es_nas

    data.frame(
      variable = variable,

      n_vacios = sum(es_vacio),
      pct_vacios = round(
        100 * sum(es_vacio) / n_total,
        2
      ),

      n_nas = sum(es_nas),
      pct_nas = round(
        100 * sum(es_nas) / n_total,
        2
      ),

      n_vacios_nas = sum(es_vacio | es_nas),
      pct_vacios_nas = round(
        100 * sum(es_vacio | es_nas) / n_total,
        2
      ),

      stringsAsFactors = FALSE
    )
  })
)

# Remove row names
rownames(resumen_ausentes) <- NULL

# Sort from highest to lowest percentage of empty cells plus NAS values
resumen_ausentes <- resumen_ausentes[
  order(
    -resumen_ausentes$pct_vacios_nas,
    resumen_ausentes$variable
  ),
]

# Open the complete table in RStudio
#View(resumen_ausentes)

```

# 1. Convert dates/times
```{r, echo=FALSE}

# Identify and report date/time values that cannot be parsed, then convert valid entries to standardized POSIXct format.
# Empty values are treated as missing (NA), while invalid formats are flagged before conversion.
# Format: POSIXct (date format)

date_vars <- intersect(c(
  "fecha_trauma", "fecha_ingreso_hospital", "fecha_ingreso_uci",
  "fecha_alta_uci", "fecha_alta_hospital", "fecha_exitus",
  "fecha_reingreso_uci", "fecha_alta_reingreso"
), names(df))

time_vars <- intersect(c(
  "hora_trauma", "hora_ingreso_hospital",
  "hora_ingreso_uci", "hora_exitus"
), names(df))

parse_columns <- function(data, vars, orders) {
  original <- data[vars]

  parsed <- lapply(original, function(x) {
    x <- trimws(as.character(x))
    x[x == ""] <- NA
    suppressWarnings(parse_date_time(x, orders = orders))
  })

  problems <- sapply(seq_along(vars), function(i) {
    x <- trimws(as.character(original[[i]]))
    sum(is.na(parsed[[i]]) & !is.na(x) & x != "")
  })

  names(problems) <- vars

  list(
    data = parsed,
    problems = problems[problems > 0]
  )
}

dates <- parse_columns(
  df,
  date_vars,
  c("dmy", "ymd", "dmy HMS", "ymd HMS")
)

times <- parse_columns(
  df,
  time_vars,
  c("HMS", "HM", "H M", "I M p")
)

df[date_vars] <- dates$data
df[time_vars] <- times$data

dates$problems
times$problems
```

# 2. Convert numeric variables
```{r, echo=FALSE}

# Clean and convert numeric variables to numeric format by trimming spaces, standardizing decimal separators, and setting invalid values to NA.
numeric_vars <- c(
  "edad",
  "creatinina_inicial", "creatinina_24h",
  "hemoglobina_inicial", "hemoglobina_24h",
  "hematocrito_inicial", "hematocrito_24h",
  "inr_inicial", "inr_24h",
  "protrombina_inicial", "protrombina_24h",
  "ttpa_inicial", "ttpa_24h",
  "fibrinogeno_inicial", "fibrinogeno_24h",
  "ph_inicial", "ph_24h",
  "exceso_base_inicial", "exceso_base_24h",
  "lactico_inicial", "lactico_24h",
  "etanol_cuantitativo_sangre",
  "fr_ph",
  "pto_fr", "pto_tas", "pto_gcs",
  "rts_t", "rts", "ps",
  "dias_uci", "dias_hospital",
  "rscore_edad", "rscore_pupilas",
  "retrascore"
)

gravedad_vars <- grep("^lesion_\\d+_gravedad$", names(df), value = TRUE)

numeric_vars <- intersect(
  unique(c(numeric_vars, gravedad_vars)),
  names(df)
)

numeric_vars <- numeric_vars[
  !vapply(df[numeric_vars], is.numeric, logical(1))
]

numeric_problems <- sapply(df[numeric_vars], function(x) {
  x2 <- trimws(as.character(x))
  x2 <- gsub(",", ".", x2)

  sum(
    is.na(suppressWarnings(as.numeric(x2))) &
      x2 != "" &
      !is.na(x2)
  )
})

numeric_problems[numeric_problems > 0]

df[numeric_vars] <- lapply(df[numeric_vars], function(x) {
  x <- trimws(as.character(x))
  x[x == ""] <- NA
  x <- gsub(",", ".", x)

  suppressWarnings(as.numeric(x))
})
```

# 3. Candidate cohort based on the RETRASCORE paper

The cohort is constructed before the subsequent exclusions based on `edad` and
`sdmo`. Following the paper, the only exclusion criterion applied here is an
unavailable in-hospital mortality outcome. Missing predictor values are retained
so they can be described and a complete-case analysis or imputation can be
chosen later.

Operationalization decisions:

- ATLS-based management is a feature of the registry protocol, and no
  patient-level variable is available for use as a filter.
- In-hospital mortality combines `evolucion_uci` and
  `evolucion_hospitalaria`, because patients who die in the ICU generally do
  not have the hospital outcome field completed.
- The analytical outcome is `mortalidad_30d`: deaths within 30 days of trauma
  are coded as 1; live hospital discharges and deaths after day 30 are coded
  as 0. Deceased patients without a valid date retain an unknown outcome.
- `probabilidad_muerte_triss` is calculated as `1 - ps/100`, because `ps` is
  stored as a survival percentage ranging from 0 to 100.
- Medical prehospital care is defined as `uvi movil` or `helicoptero`; the
  categories `no medicalizada`, `sin atencion prehospitalaria`, and
  `soporte vital enfermeria` form the reference group.
- Hemodynamic failure is approximated by `shock` or `shock refractario` in
  `situacion_hemodinamica`; respiratory failure is defined as PaO2/FiO2 less
  than or equal to 200 in `adrs`.
- `situacion_hemodinamica`, `adrs`, `coagulopatia`, `vmec_dico`, and
  `hemorragia_masiva` are assumed to represent the first 24 hours, as defined
  in the registry. The export has no timestamp that allows this to be verified
  for each patient.
- mais1-mais6 are used directly according to the current code mapping; their
  anatomical equivalence and the interpretation of empty cells require
  confirmation against the registry dictionary, not only percentage comparisons.

```{r}
# Helper functions ---------------------------------------------------------

limpiar_texto <- function(x) {
  x <- tolower(trimws(as.character(x)))
  x[x %in% c("", "na", "nas")] <- NA_character_
  x
}

convertir_numero <- function(x) {
  x <- trimws(as.character(x))
  x[x %in% c("", "NA", "NAS")] <- NA_character_
  suppressWarnings(as.numeric(gsub(",", ".", x)))
}

# In-hospital outcome. Patients who die in the ICU generally have an empty
# hospital outcome, so both outcome fields are combined.
evolucion_uci_limpia <- limpiar_texto(df$evolucion_uci)
evolucion_hospital_limpia <- limpiar_texto(df$evolucion_hospitalaria)

mortalidad_intrahospitalaria <- case_when(
  evolucion_uci_limpia %in% c("exitus", "exitus y donacion") ~ 1L,
  evolucion_hospital_limpia %in% c(
    "exitus", "exitus y donacion"
  ) ~ 1L,
  evolucion_hospital_limpia == "alta" ~ 0L,
  TRUE ~ NA_integer_
)

# Primary outcome from the paper: mortality within 30 days of trauma.
# fecha_alta_uci is used for ICU deaths and fecha_alta_hospital for deaths
# after ICU discharge. fecha_exitus is used as a fallback.
muerte_uci <- evolucion_uci_limpia %in% c(
  "exitus", "exitus y donacion"
)
muerte_hospital <- evolucion_hospital_limpia %in% c(
  "exitus", "exitus y donacion"
)

fecha_muerte <- as.POSIXct(
  rep(NA_real_, nrow(df)),
  origin = "1970-01-01",
  tz = "UTC"
)
fecha_muerte[muerte_uci] <- df$fecha_alta_uci[muerte_uci]
fecha_muerte[muerte_hospital] <- df$fecha_alta_hospital[
  muerte_hospital
]

usar_fecha_exitus <- mortalidad_intrahospitalaria == 1 &
  is.na(fecha_muerte) & !is.na(df$fecha_exitus)
usar_fecha_exitus[is.na(usar_fecha_exitus)] <- FALSE
fecha_muerte[usar_fecha_exitus] <- df$fecha_exitus[usar_fecha_exitus]

dias_hasta_muerte <- as.numeric(
  difftime(fecha_muerte, df$fecha_trauma, units = "days")
)
fecha_muerte_valida <- !is.na(dias_hasta_muerte) & dias_hasta_muerte >= 0

# Survival beyond day 30 is considered 30-day survival even if the patient
# dies later; these later deaths are not excluded.
mortalidad_30d <- case_when(
  mortalidad_intrahospitalaria == 0 ~ 0L,
  mortalidad_intrahospitalaria == 1 &
    fecha_muerte_valida & dias_hasta_muerte <= 30 ~ 1L,
  mortalidad_intrahospitalaria == 1 &
    fecha_muerte_valida & dias_hasta_muerte > 30 ~ 0L,
  TRUE ~ NA_integer_
)

# Comparators used in the paper.
dias_estancia_uci <- convertir_numero(df$dias_uci)
ps_triss_pct <- convertir_numero(df$ps)
probabilidad_muerte_triss <- ifelse(
  !is.na(ps_triss_pct) & ps_triss_pct >= 0 & ps_triss_pct <= 100,
  1 - ps_triss_pct / 100,
  NA_real_
)

# Patient-related variables -----------------------------------------------

genero_limpio <- limpiar_texto(df$genero)
genero_binario <- case_when(
  genero_limpio == "masculino" ~ 1L,
  genero_limpio == "femenino" ~ 0L,
  TRUE ~ NA_integer_
)

edad_numerica <- convertir_numero(df$edad)
edad_categoria <- cut(
  edad_numerica,
  breaks = c(-Inf, 49, 65, 75, Inf),
  labels = c("<50", "50-65", "66-75", ">75"),
  right = TRUE
)

antiagregado_limpio <- limpiar_texto(df$antiagregado)
dicumarinicos_limpio <- limpiar_texto(df$dicumarinicos)
acod_limpio <- limpiar_texto(df$acod)

precoag <- case_when(
  antiagregado_limpio == "si" |
    dicumarinicos_limpio == "si" |
    acod_limpio == "si" ~ 1L,
  antiagregado_limpio == "no" &
    dicumarinicos_limpio == "no" &
    acod_limpio == "no" ~ 0L,
  TRUE ~ NA_integer_
)

# Prehospital variables ----------------------------------------------------

atencion_ph_limpia <- limpiar_texto(df$atencion_prehospitalaria)
atencion_prehospitalaria_medica <- case_when(
  atencion_ph_limpia %in% c("uvi movil", "helicoptero") ~ 1L,
  atencion_ph_limpia %in% c(
    "no medicalizada", "sin atencion prehospitalaria",
    "soporte vital enfermeria"
  ) ~ 0L,
  TRUE ~ NA_integer_
)

iot_ph_limpia <- limpiar_texto(df$iot_prehospitalaria)
iot_prehospitalaria_binaria <- case_when(
  iot_ph_limpia %in% c(
    "si", "si, con empleo de dispositivos de apoyo"
  ) ~ 1L,
  iot_ph_limpia == "no" ~ 0L,
  TRUE ~ NA_integer_
)

tipo_trauma_limpio <- limpiar_texto(df$tipo_trauma)
trauma_penetrante <- case_when(
  tipo_trauma_limpio == "penetrante" ~ 1L,
  tipo_trauma_limpio == "contuso" ~ 0L,
  TRUE ~ NA_integer_
)

mecanismo_limpio <- limpiar_texto(df$mecanismo)
# Requested grouping: other and unknown form a single category.
# Empty values remain NA and are not included in this category.
mecanismo_limpio[mecanismo_limpio %in% c(
  "otro", "otros", "desconocido"
)] <- "otros/desconocido"
intencionalidad_limpia <- limpiar_texto(df$intencionalidad)

tipo_lesion_categoria <- factor(
  case_when(
    intencionalidad_limpia == "no intencional : trafico" ~
      "Road Traffic accident",
    intencionalidad_limpia == "no intencional : caida" ~ "Fall",
    intencionalidad_limpia == "no intencional : laboral" ~
      "Occupational accident",
    intencionalidad_limpia == "no intencional : deportivo" ~
      "Sport related",
    intencionalidad_limpia == "intencionado : agresion" ~ "Assault",
    intencionalidad_limpia == "intencionado : autolisis" ~ "Self-injury",
    intencionalidad_limpia == "desconocido" ~ "Unknown",
    intencionalidad_limpia == "no intencional : otras actividades" ~
      "Other",
    TRUE ~ NA_character_
  ),
  levels = c(
    "Road Traffic accident", "Fall", "Occupational accident",
    "Sport related", "Assault", "Self-injury", "Unknown", "Other"
  )
)

mecanismo_alto_riesgo <- case_when(
  mecanismo_limpio %in% c(
    "arma de fuego", "atropello", "caida accidental", "otros/desconocido"
  ) ~ 1L,
  mecanismo_limpio == "precipitacion" &
    intencionalidad_limpia == "intencionado : autolisis" ~ 1L,
  !is.na(mecanismo_limpio) & mecanismo_limpio != "precipitacion" ~ 0L,
  mecanismo_limpio == "precipitacion" &
    !is.na(intencionalidad_limpia) ~ 0L,
  TRUE ~ NA_integer_
)

# Physiological variables --------------------------------------------------

pupilas_limpias <- limpiar_texto(df$pupilas_ingreso)
pupilas_categoria <- factor(
  case_when(
    pupilas_limpias == "sin alteraciones" ~ "normal",
    pupilas_limpias == "midriasis unilateral" ~ "midriasis unilateral",
    pupilas_limpias == "midriasis bilatera" ~ "midriasis bilateral",
    TRUE ~ NA_character_
  ),
  levels = c("normal", "midriasis unilateral", "midriasis bilateral")
)

gcs_total <- convertir_numero(df$gcs_total_introducido)
gcs_menor_igual_8 <- ifelse(
  is.na(gcs_total), NA_integer_, as.integer(gcs_total <= 8)
)

# Anatomical variables -----------------------------------------------------
# RETRAUCI regional MAIS variables are used directly:
# mais1 head, mais2 upper extremity, mais3 thorax, mais4 abdomen,
# mais5 lower extremity, and mais6 external/thermal injuries.
# In this export, an empty MAIS cell means no injury was documented in that
# region and is coded as 0. Non-empty non-numeric values or values outside the
# AIS 0-6 range are retained as missing.

convertir_mais_regional <- function(x) {
  texto <- trimws(as.character(x))
  es_vacio <- is.na(texto) | texto == "" |
    toupper(texto) %in% c("NA", "NAS")
  resultado <- suppressWarnings(as.numeric(gsub(",", ".", texto)))
  resultado[es_vacio] <- 0
  resultado[!is.na(resultado) & !(resultado %in% 0:6)] <- NA_real_
  resultado
}

dicotomizar_mais <- function(x) {
  ifelse(is.na(x), NA_integer_, as.integer(x >= 3))
}

ais_cabeza <- convertir_mais_regional(df$mais1)
ais_ext_superior <- convertir_mais_regional(df$mais2)
ais_torax <- convertir_mais_regional(df$mais3)
ais_abdomen <- convertir_mais_regional(df$mais4)
ais_ext_inferior <- convertir_mais_regional(df$mais5)
ais_externo <- convertir_mais_regional(df$mais6)

# Organ failures within the first 24 hours --------------------------------

hemodinamica_limpia <- limpiar_texto(df$situacion_hemodinamica)
fallo_hemodinamico <- case_when(
  hemodinamica_limpia %in% c("shock", "shock refractario") ~ 1L,
  hemodinamica_limpia %in% c(
    "estable", "inestable remonta con volumen"
  ) ~ 0L,
  TRUE ~ NA_integer_
)

respiratorio_limpio <- limpiar_texto(df$adrs)
fallo_respiratorio <- case_when(
  respiratorio_limpio %in% c("<100", "101-200") ~ 1L,
  respiratorio_limpio %in% c("no", "201 -300", "> 300") ~ 0L,
  TRUE ~ NA_integer_
)

coagulopatia_limpia <- limpiar_texto(df$coagulopatia)
coagulopatia_24h <- case_when(
  coagulopatia_limpia == "si" ~ 1L,
  coagulopatia_limpia == "no" ~ 0L,
  TRUE ~ NA_integer_
)

# Treatments within the first 24 hours ------------------------------------

vm_limpia <- limpiar_texto(df$vmec_dico)
ventilacion_mecanica <- case_when(
  vm_limpia == "si" ~ 1L,
  vm_limpia == "no" ~ 0L,
  TRUE ~ NA_integer_
)

hemorragia_limpia <- limpiar_texto(df$hemorragia_masiva)
hemorragia_masiva <- case_when(
  hemorragia_limpia == "si" ~ 1L,
  hemorragia_limpia == "no" ~ 0L,
  TRUE ~ NA_integer_
)

# Final cohort -------------------------------------------------------------

cohorte_retrascore <- data.frame(
  id_fila = seq_len(nrow(df)),
  historia = df$historia,
  registro = df$registro,
  hospital = df$hospital,
  año_trauma = year(df$fecha_trauma),
  mortalidad_intrahospitalaria = mortalidad_intrahospitalaria,
  mortalidad_30d = mortalidad_30d,
  dias_hasta_muerte = dias_hasta_muerte,
  dias_estancia_uci = dias_estancia_uci,
  probabilidad_muerte_triss = probabilidad_muerte_triss,
  genero_masculino = genero_binario,
  edad = edad_numerica,
  edad_categoria = edad_categoria,
  precoag = precoag,
  atencion_prehospitalaria_medica = atencion_prehospitalaria_medica,
  iot_prehospitalaria = iot_prehospitalaria_binaria,
  trauma_penetrante = trauma_penetrante,
  tipo_lesion = tipo_lesion_categoria,
  mecanismo_agrupado = mecanismo_limpio,
  mecanismo_alto_riesgo = mecanismo_alto_riesgo,
  pupilas = pupilas_categoria,
  gcs_total = gcs_total,
  gcs_menor_igual_8 = gcs_menor_igual_8,
  ais_cabeza = ais_cabeza,
  mais_cabeza = dicotomizar_mais(ais_cabeza),
  ais_torax = ais_torax,
  mais_torax = dicotomizar_mais(ais_torax),
  ais_abdomen = ais_abdomen,
  mais_abdomen = dicotomizar_mais(ais_abdomen),
  ais_ext_superior = ais_ext_superior,
  mais_ext_superior = dicotomizar_mais(ais_ext_superior),
  ais_ext_inferior = ais_ext_inferior,
  mais_ext_inferior = dicotomizar_mais(ais_ext_inferior),
  ais_externo = ais_externo,
  mais_externo = dicotomizar_mais(ais_externo),
  fallo_hemodinamico = fallo_hemodinamico,
  fallo_respiratorio = fallo_respiratorio,
  coagulopatia_24h = coagulopatia_24h,
  ventilacion_mecanica = ventilacion_mecanica,
  hemorragia_masiva = hemorragia_masiva
)

n_inicial <- nrow(cohorte_retrascore)
n_sin_mortalidad <- sum(
  is.na(cohorte_retrascore$mortalidad_intrahospitalaria)
)

cohorte_retrascore <- cohorte_retrascore %>%
  filter(!is.na(mortalidad_intrahospitalaria))

flujo_cohorte_retrascore <- data.frame(
  etapa = c(
    "Registros iniciales",
    "Excluidos: mortalidad intrahospitalaria desconocida",
    "Cohorte RETRASCORE"
  ),
  n = c(n_inicial, n_sin_mortalidad, nrow(cohorte_retrascore))
)

flujo_cohorte_retrascore
table(cohorte_retrascore$mortalidad_30d, useNA = "ifany")

resumen_outcome_30d <- data.frame(
  medida = c(
    "Cohorte elegible",
    "Outcome a 30 dias evaluable",
    "Muertes hasta 30 dias",
    "Muertes intrahospitalarias posteriores a 30 dias",
    "Muertes sin fecha valida para determinar 30 dias"
  ),
  n = c(
    nrow(cohorte_retrascore),
    sum(!is.na(cohorte_retrascore$mortalidad_30d)),
    sum(cohorte_retrascore$mortalidad_30d == 1, na.rm = TRUE),
    sum(
      cohorte_retrascore$mortalidad_intrahospitalaria == 1 &
        cohorte_retrascore$dias_hasta_muerte > 30,
      na.rm = TRUE
    ),
    sum(
      cohorte_retrascore$mortalidad_intrahospitalaria == 1 &
        is.na(cohorte_retrascore$mortalidad_30d)
    )
  )
)

resumen_outcome_30d

# Availability of all candidate variables in the cohort.
resumen_ausentes_cohorte <- data.frame(
  variable = names(cohorte_retrascore),
  n_ausentes = vapply(cohorte_retrascore, function(x) {
    sum(is.na(x))
  }, numeric(1)),
  pct_ausentes = round(vapply(cohorte_retrascore, function(x) {
    100 * mean(is.na(x))
  }, numeric(1)), 2),
  row.names = NULL
)

resumen_ausentes_cohorte <- resumen_ausentes_cohorte[
  order(-resumen_ausentes_cohorte$pct_ausentes),
]
resumen_ausentes_cohorte
```

# 4. Save the preprocessed cohort

The generated file is the interface with the model development and validation
document. It must be regenerated whenever the source data or preprocessing
rules change.

```{r}
saveRDS(cohorte_retrascore, "cohorte_retrascore.rds")
```
